Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-09-25 arxiv/proof_examples.tex
reorg
fix links

\section{Proof of the Special Case}\label{app_iso_cov}
\todo{switch the $a_3$ and $a_4$}

@@

For any fixed $v\in \R$, $\beta_1 - v\beta_2$ is a random vector with i.i.d. entries with mean $0$ and variance $(1-v)^2\kappa^2 + d^2$. Then using the concentration result, Lemma \ref{largedeviation}, we get that for any constant $\e>0$, 
\begin{align}
&\left|\delta_{\bias}(v)- [(1-v)^2\kappa^2 + d^2]\tr (\cal K^\top \cal K)\right| \nonumber\\
&=  \left|(\beta_1 - {v} \beta_2)^\top \cal K^\top \cal K (\beta_1 - {v} \beta_2) - [(1-v)^2\kappa^2 + d^2]\tr (\cal K^\top \cal K)\right| \nonumber\\
&\le p^\e [(1-v)^2\kappa^2 + d^2] \left\{\tr\left[(\cal K^\top \cal K)^2\right]\right\}^{1/2} \lesssim p^{1/2+\e} [(1-v)^2\kappa^2 + d^2] \quad \text{w.h.p.},\label{anotherbeta}
\end{align}
where $\delta_{\bias}(v)$ was defined in \eqref{revise_deltabias}, $\cal K$ was defined as $\cal K:=v\Sigma_2^{1/2}({v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} X_1^{\top}X_1$, and in the last step we used $\|\cal K\|=\OO(1)$ by \eqref{eq_isometric}. Now for $\tr (\cal K^\top \cal K)$, we rewrite it as
\begin{align*}
v^2 \tr \left[ ({v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2({v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}(X_1^\top X_1)^2 \right].
\end{align*}
Bounding $(X_1^\top X_1)^2=(Z_1^\top Z_1)^2$ using \eqref{eq_isometric}, we obtain that
\be\label{anotherwtbeta} \delta^\star_{\bias}(v) \cdot (\al_-^2(\rho_1)-\OO(p^{-1/2+\e})) \le  [(1-v)^2\kappa^2 + d^2]\tr (\cal K^\top \cal K) \le \delta^\star_{\bias}(v) \cdot  (\al_+^2(\rho_1)+\OO(p^{-1/2+\e})),\ee
where
$$\delta^\star_{\bias}(v):=n_1^2 v^2 [(1-v)^2\kappa^2 + d^2]\tr \left[ ({v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2({v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\right].$$
Note that $\delta^\star_{\bias}(v) \sim p$ by \eqref{eq_isometric}. Hence combining \eqref{anotherbeta} and \eqref{anotherwtbeta} we get
\be\label{replaceest}
  \delta^\star_{\bias}(v) \cdot (\al_-^2(\rho_1)-\OO(p^{-1/2+\e})) \le  \delta_{\bias}(v)\le \delta^\star_{\bias}(v) \cdot  (\al_+^2(\rho_1)+\OO(p^{-1/2+\e})).
\ee
Now we can replace the estimate \eqref{bounddelta-} with this stronger estimate, and repeat all the other parts of the proof of Lemma \ref{thm_model_shift} to conclude Lemma \ref{prop_model_shift_tight}. In particular, one can calculate $  \delta^\star_{\bias}(v)$ using Lemma \ref{lem_cov_derivative} and get the $  \Delta^\star_{\bias}(v)$ term. 
\end{proof}

@@

\begin{proof}
	Using Lemma \ref{largedeviation} again, we can simplify \eqref{revise_eq_val_mtl} as $\val(v)= N_2h(v) \cdot  \left( 1+\OO(p^{-1/2+\e})\right)$, where the function $ h$ is defined as
	%We define the function
	\begin{align}
		h(v) =& \frac{\rho_1}{\rho_2}\left[d^2 +\left( v-1\right)^2\kappa^2\right]\cdot \tr\left[( v^2X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_2^{\top}X_2)^2\right] \nonumber \\
		& +  v^2\left[d^2 +\left( v-1\right)^2\kappa^2\right]\cdot \tr\left[( v^2X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \nonumber\\
		& + \left(\frac{\rho_1}{\rho_2} v^2 + 1\right)\sigma^2 \cdot \bigtr{(v^2X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }. \nonumber
	\end{align}
	%\cor Under the setting of Lemma \ref{prop_model_shift_tight},
	Furthermore, the validation loss in \eqref{approxvalid} reduces to
	\be\label{boundv-w}
		g(v)=\left[N_2 h(v) + (N_1+N_2)\sigma^2\right]\cdot \left( 1+\OO(p^{-(1-\e_0)/2+\e})\right)\quad \text{w.h.p.}
	\ee
	for any constant $\e>0$. %\nc Thus for the following discussions, it suffices to focus on the behavior of $h (v)$.
Let $\hat w$ be the minimizer of $h(v)$. Our proof consists of the following two steps.
\squishlist
	\item First, we show that $\hat{w}$ is close to $1$.
	\item Second, with \eqref{boundv-w} we show that $\hat v$ is close to $\hat w$. Then we plug $\hat{v}$ into $\te(\hat{\beta}_2^{\MTL})$ to get \eqref{simple1}.
\squishend
%For the minimizer $\hat w$ of $\val(w)$, we have a similar result as in Proposition \ref{thm_cov_shift}.
For the first step, we will prove the following result.
\begin{claim}\label{lem_hat_v}
%Suppose the assumptions of Lemma \ref{prop_model_shift_tight} hold. Assume that $ \kappa^2 \sim pd^2 \sim \sigma^2$ are of the same order.
There exists a constant $C>0$ such that
	\be\label{hatw_add1}|\hat w -1|\le C\left(\frac{d^2}{\kappa^2} + \frac{\sigma^2}{p\kappa^2}\right)\quad \text{w.h.p.}\ee
	%is $\hat{w} = 1 \pm \bigo{\frac 1 {n_1+n_2}}$ \todo{(figure out the constants)}
\end{claim}
\begin{proof}
To be consistent with the notation $\hat w$, we change the name of the argument to $w$ in the proof. First it is easy to observe that $h(w)< h(-w)$ for $w> 0$. Second, using \eqref{eq_isometric}, it is easy to check that $h(1)=\OO(pd^2 + \sigma^2)$ and $ h(w)\gtrsim p\kappa^2 \gg h(1)$ for $w\ge 2$ or $w\le 1/2$.  Hence it suffices to consider the case $1/2\le w\le 2$.

We first consider the case $1\le w\le 2$. We write
\begin{align}
	h(w) =&  \frac{\rho_1}{\rho_2}\left[\frac{d^2}{w^4} +\frac{\left( w-1\right)^2}{w^4}\kappa^2\right]\cdot \tr\left[( X_1^{\top}X_1 +w^{-2}X_2^{\top}X_2)^{-2} (X_2^{\top}X_2)^2\right] \nonumber \\
	& + \left[\frac{d^2}{w^2} +\frac{\left( w-1\right)^2}{w^2}\kappa^2\right]\cdot \tr\left[( X_1^{\top}X_1 +w^{-2}X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \nonumber\\
			& +  \frac{\rho_1}{\rho_2}\sigma^2  \cdot \bigtr{(X_1^{\top}X_1  + w^{-2}X_2^{\top}X_2)^{-1} }+ \sigma^2 \cdot \bigtr{(w^2X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }.\nonumber%+  \sigma^2n_2 \cdot \bigtr{(X_1^{\top}X_1  + w^{-2}X_2^{\top}X_2)^{-1} }. \nonumber
\end{align}
Notice that
$$\tr\left[( X_1^{\top}X_1 +w^{-2}X_2^{\top}X_2)^{-2} (X_i^{\top}X_i)^2\right],\ \ i=1,2, \quad \text{and} \quad \bigtr{(X_1^{\top}X_1  + w^{-2}X_2^{\top}X_2)^{-1} }$$
are increasing functions in $w$. Hence taking derivative of $h(w)$ with respect to $w$, we obtain that
\begin{align*}
h'(w) \ge & \frac{\rho_1}{\rho_2}\left[ \frac{2(w-1)(2-w)}{w^5}\kappa^2 - \frac{4d^2}{w^5}\right] \tr\left[( X_1^{\top}X_1 +w^{-2}X_2^{\top}X_2)^{-2} (X_2^{\top}X_2)^2\right]   \\
 &+ \left[\frac{2\left( w-1\right)}{w^3}\kappa^2 - \frac{2d^2}{w^3} \right]\cdot \tr\left[( X_1^{\top}X_1 +w^{-2}X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \nonumber\\
		&- 2  \frac{\sigma^2}{w^3} \cdot \bigtr{(X_1^{\top}X_1 + w^{-2}X_2^{\top}X_2)^{-2} X_1^\top X_1  } =  \bigtr{(X_1^{\top}X_1 + w^{-2}X_2^{\top}X_2)^{-2} \cal A },
\end{align*}
where the matrix $\cal A$ is defined as 
\begin{align*}
\cal A :&= \frac{\rho_1}{\rho_2}\left[ \frac{2(w-1)(2-w)}{w^5}\kappa^2 - \frac{4d^2}{w^5}\right](X_2^{\top}X_2)^2 + \left[\frac{2\left( w-1\right)}{w^3}\kappa^2 - \frac{2d^2}{w^3} \right](X_1^{\top}X_1)^2 - 2 \frac{\sigma^2}{w^3}X_1^\top X_1.
\end{align*}
Using the estimate \eqref{eq_isometric}, we get that $\cal A$ is lower bounded as
\begin{align*}
\cal A \succeq&~ - \frac{4d^2}{w^5}n_1n_2 (\al_+(\rho_2)+\oo(1))^2 + \left[\frac{2\left( w-1\right)}{w^3}\kappa^2 - \frac{2d^2}{w^3} \right]n_1^2 (\al_-(\rho_1)-\oo(1))^2 \\
&~ - 2 \frac{\sigma^2}{w^3}n_1(\al_{+}(\rho_1)+\oo(1)) \succ 0,
\end{align*}
as long as
$$w> w_1:=1 +\frac{d^2}{\kappa^2}+ \frac{\sigma^2}{n_1\kappa^2}\frac{\al_{+}(\rho_1)+\oo(1)}{\al_{-}^2(\rho_1)} + \frac{2d^2}{\kappa^2}\frac{\rho_2(\al_+^2(\rho_2)+\oo(1))}{\rho_1\al_-^2(\rho_1) }.$$
Hence $h'(w)>0$ on $(w_1,\infty)$, %i.e. $h(w)$ is strictly increasing for $w>w_1$. This 
which gives $\hat w\le w_1$.

For the case $1/2\le w\le 1$, the proof is similar as above. Taking derivative of $h(w)$, we obtain that
\begin{align}
	h'(w) \le& \frac{\rho_1}{\rho_2} \left[2\left( w-1\right) \kappa^2\right]\cdot \tr\left[( w^2X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_2^{\top}X_2)^2\right] \nonumber \\
	& +  \left[2wd^2 +2w\left( w-1\right)(2w-1)\kappa^2\right]\cdot \tr\left[( w^2X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \nonumber\\
	& + \frac{\rho_1}{\rho_2}(2 w\sigma^2) \cdot \bigtr{(w^2X_1^{\top}X_1 + X_2^{\top}X_2)^{-2} X_2^\top X_2  }= \frac{\rho_1}{\rho_2} \bigtr{(w^2X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} \cal B} , \nonumber
\end{align}
where the matrix $\cal B$ is
$$\cal B= 2\left( w-1\right) \kappa^2  (X_2^{\top}X_2)^2+\frac{\rho_2}{\rho_1}\left[2wd^2 +2w\left( w-1\right)(2w-1)\kappa^2\right](X_1^{\top}X_1)^2 + 2 w\sigma^2 X_2^\top X_2 .$$
Using the estimate \eqref{eq_isometric}, we get that $\cal B$ is upper bounded as
\begin{align*}
\cal B \preceq - 2(1-w)\kappa^2 n_2^2 (\al_-(\rho_2) -\oo(1))^2 +2w d^2 n_1n_2 (\al_+(\rho_1) +\oo(1))^2 + 2w\sigma^2 n_2 (\al_+(\rho_2)+\oo(1)) \prec 0,
\end{align*}
as long as
$$w< w_2:=1 -   \frac{d^2}{\kappa^2}\frac{\rho_1(\al_+(\rho_1) +\oo(1))^2}{\rho_2 \al_-^2(\rho_2) } -  \frac{\sigma^2}{n_2\kappa^2}\frac{\al_{+}(\rho_2)+\oo(1)}{\al_{-}^2(\rho_2)} .$$
Hence $h'(w)<0$ on $[0,w_2)$, %i.e. $h(w)$ is strictly decreasing for $w<w_2$. This 
which gives $\hat w\ge w_2$.

@@

Inserting \eqref{hatv_add1} into \eqref{eq_te_mtl_2task} and applying Lemma  \ref{largedeviation} to $(\beta_1-\hat v\beta_s)$, we get that w.h.p.,
\begin{align}
\te(\hat{\beta}_2^{\MTL})&=(1+\OO(\cal E))\cdot \left[d^2 + \OO\left(\cal E^2 \kappa^2\right)\right] \tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right]\nonumber\\
&+(1+\OO(\cal E))\cdot \sigma^2  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }. \label{1Liso}
\end{align}
In order to study the phenomenon of bias-variance trade-off, we need the bias term with $d^2$ and the variance term with $\sigma^2$ to be of the same order. With \eqref{eq_isometric}, we have that
$$\tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \sim p,\quad \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} } \sim 1.$$
Hence we need to choose that $p d^2 \sim \sigma^2$. On the other hand, we want the error term $\cal E^2 \kappa^2$ to be much smaller than $d^2$, which leads to the condition $p^{-1+\e_0+4\e}\kappa^2  \ll d^2 \ll \kappa^2$. The above considerations lead to the choices of parameters in \eqref{choiceofpara}, under which we can simplify \eqref{1Liso} to \eqref{simple1}.
%\begin{align}
%\te(\hat{\beta}_2^{\MTL})&=(1+\OO(n^{-\e}))\cdot d^2 \tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \nonumber \\
%&+(1+\OO(n^{-\e}))\cdot \sigma^2  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }  \label{simple1}
%\end{align}
%w.h.p. for some constant $\e>0$.
\end{proof}

@@

\subsection{Task Similarity}\label{app_proof_31}

@@

>>>>

\subsection{Proof of the Special Case}\label{app_iso_cov}
\todo{switch the $a_3$ and $a_4$}

@@

For any fixed $v\in \R$, $\beta_1 - v\beta_2$ is a random vector with i.i.d. entries with mean $0$ and variance $(1-v)^2\kappa^2 + d^2$. Then using the concentration result, Lemma \ref{largedeviation}, we get that for any constant $\e>0$, 
\begin{align}
&\left|\delta_{\bias}(v)- [(1-v)^2\kappa^2 + d^2]\tr (\cal K^\top \cal K)\right| \nonumber\\
&=  \left|(\beta_1 - {v} \beta_2)^\top \cal K^\top \cal K (\beta_1 - {v} \beta_2) - [(1-v)^2\kappa^2 + d^2]\tr (\cal K^\top \cal K)\right| \nonumber\\
&\le p^\e [(1-v)^2\kappa^2 + d^2] \left\{\tr\left[(\cal K^\top \cal K)^2\right]\right\}^{1/2} \lesssim p^{1/2+\e} [(1-v)^2\kappa^2 + d^2] \quad \text{w.h.p.},\label{anotherbeta}
\end{align}
where $\delta_{\bias}(v)$ was defined in \eqref{revise_deltabias}, $\cal K$ was defined as $\cal K:=v\Sigma_2^{1/2}({v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} X_1^{\top}X_1$, and in the last step we used $\|\cal K\|=\OO(1)$ by Fact \ref{lem_minv}. Now for $\tr (\cal K^\top \cal K)$, we rewrite it as
\begin{align*}
v^2 \tr \left[ ({v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2({v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}(X_1^\top X_1)^2 \right].
\end{align*}
Bounding $(X_1^\top X_1)^2=(Z_1^\top Z_1)^2$ using Fact \eqref{lem_minv}, we obtain that
\be\label{anotherwtbeta} \delta^\star_{\bias}(v) \cdot (\al_-^2(\rho_1)-\OO(p^{-1/2+\e})) \le  [(1-v)^2\kappa^2 + d^2]\tr (\cal K^\top \cal K) \le \delta^\star_{\bias}(v) \cdot  (\al_+^2(\rho_1)+\OO(p^{-1/2+\e})),\ee
where
$$\delta^\star_{\bias}(v):=n_1^2 v^2 [(1-v)^2\kappa^2 + d^2]\tr \left[ ({v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2({v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\right].$$
Note that $\delta^\star_{\bias}(v) \sim p$ by Fact \ref{lem_minv}. Hence combining \eqref{anotherbeta} and \eqref{anotherwtbeta} we get
\be\label{replaceest}
  \delta^\star_{\bias}(v) \cdot (\al_-^2(\rho_1)-\OO(p^{-1/2+\e})) \le  \delta_{\bias}(v)\le \delta^\star_{\bias}(v) \cdot  (\al_+^2(\rho_1)+\OO(p^{-1/2+\e})).
\ee
Now we can replace the estimate \eqref{bounddelta-} with this stronger estimate, and repeat all the other parts of the proof of Lemma \ref{thm_model_shift} to conclude Lemma \ref{prop_model_shift_tight}. In particular, one can calculate $  \delta^\star_{\bias}(v)$ using Lemma \ref{lem_cov_derivative} and get the $  \Delta^\star_{\bias}(v)$ term. 
\end{proof}

@@

\begin{proof}
	Using Lemma \ref{largedeviation} again, we can simplify \eqref{revise_eq_val_mtl} as $\val(v)= N_2h(v) \cdot  \left( 1+\OO(p^{-1/2+\e})\right)$, where the function $ h$ is defined as
	%We define the function
	\begin{align}
		h(v) =& \frac{\rho_1}{\rho_2}\left[d^2 +\left( v-1\right)^2\kappa^2\right]\cdot \tr\left[( v^2X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_2^{\top}X_2)^2\right] \nonumber \\
		& +  v^2\left[d^2 +\left( v-1\right)^2\kappa^2\right]\cdot \tr\left[( v^2X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \nonumber\\
		& + \left(\frac{\rho_1}{\rho_2} v^2 + 1\right)\sigma^2 \cdot \bigtr{(v^2X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }. \nonumber
	\end{align}
	%\cor Under the setting of Lemma \ref{prop_model_shift_tight},
	Furthermore, the validation loss in \eqref{approxvalid} reduces to
	\be\label{boundv-w}
		g(v)=\left[N_2 h(v) + (N_1+N_2)\sigma^2\right]\cdot \left( 1+\OO(p^{-(1-\e_0)/2+\e})\right)\quad \text{w.h.p.}
	\ee
	for any constant $\e>0$. %\nc Thus for the following discussions, it suffices to focus on the behavior of $h (v)$.
Let $\hat w$ be the minimizer of $h(v)$. Our proof consists of the following two steps.
\squishlist
	\item First, we show that $\hat{w}$ is close to $1$.
	\item Second, with \eqref{boundv-w} we show that $\hat v$ is close to $\hat w$. Then we plug $\hat{v}$ into $\te(\hat{\beta}_2^{\MTL})$ to get \eqref{simple1}.
\squishend
%For the minimizer $\hat w$ of $\val(w)$, we have a similar result as in Proposition \ref{thm_cov_shift}.
For the first step, we will prove the following result.
\begin{claim}\label{lem_hat_v}
%Suppose the assumptions of Lemma \ref{prop_model_shift_tight} hold. Assume that $ \kappa^2 \sim pd^2 \sim \sigma^2$ are of the same order.
There exists a constant $C>0$ such that
	\be\label{hatw_add1}|\hat w -1|\le C\left(\frac{d^2}{\kappa^2} + \frac{\sigma^2}{p\kappa^2}\right)\quad \text{w.h.p.}\ee
	%is $\hat{w} = 1 \pm \bigo{\frac 1 {n_1+n_2}}$ \todo{(figure out the constants)}
\end{claim}
\begin{proof}
To be consistent with the notation $\hat w$, we change the name of the argument to $w$ in the proof. First it is easy to observe that $h(w)< h(-w)$ for $w> 0$. Second, using Fact \eqref{lem_minv}, it is easy to check that $h(1)=\OO(pd^2 + \sigma^2)$ and $ h(w)\gtrsim p\kappa^2 \gg h(1)$ for $w\ge 2$ or $w\le 1/2$.  Hence it suffices to consider the case $1/2\le w\le 2$.

We first consider the case $1\le w\le 2$. We write
\begin{align}
	h(w) =&  \frac{\rho_1}{\rho_2}\left[\frac{d^2}{w^4} +\frac{\left( w-1\right)^2}{w^4}\kappa^2\right]\cdot \tr\left[( X_1^{\top}X_1 +w^{-2}X_2^{\top}X_2)^{-2} (X_2^{\top}X_2)^2\right] \nonumber \\
	& + \left[\frac{d^2}{w^2} +\frac{\left( w-1\right)^2}{w^2}\kappa^2\right]\cdot \tr\left[( X_1^{\top}X_1 +w^{-2}X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \nonumber\\
			& +  \frac{\rho_1}{\rho_2}\sigma^2  \cdot \bigtr{(X_1^{\top}X_1  + w^{-2}X_2^{\top}X_2)^{-1} }+ \sigma^2 \cdot \bigtr{(w^2X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }.\nonumber%+  \sigma^2n_2 \cdot \bigtr{(X_1^{\top}X_1  + w^{-2}X_2^{\top}X_2)^{-1} }. \nonumber
\end{align}
Notice that
$$\tr\left[( X_1^{\top}X_1 +w^{-2}X_2^{\top}X_2)^{-2} (X_i^{\top}X_i)^2\right],\ \ i=1,2, \quad \text{and} \quad \bigtr{(X_1^{\top}X_1  + w^{-2}X_2^{\top}X_2)^{-1} }$$
are increasing functions in $w$. Hence taking derivative of $h(w)$ with respect to $w$, we obtain that
\begin{align*}
h'(w) \ge & \frac{\rho_1}{\rho_2}\left[ \frac{2(w-1)(2-w)}{w^5}\kappa^2 - \frac{4d^2}{w^5}\right] \tr\left[( X_1^{\top}X_1 +w^{-2}X_2^{\top}X_2)^{-2} (X_2^{\top}X_2)^2\right]   \\
 &+ \left[\frac{2\left( w-1\right)}{w^3}\kappa^2 - \frac{2d^2}{w^3} \right]\cdot \tr\left[( X_1^{\top}X_1 +w^{-2}X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \nonumber\\
		&- 2  \frac{\sigma^2}{w^3} \cdot \bigtr{(X_1^{\top}X_1 + w^{-2}X_2^{\top}X_2)^{-2} X_1^\top X_1  } =  \bigtr{(X_1^{\top}X_1 + w^{-2}X_2^{\top}X_2)^{-2} \cal A },
\end{align*}
where the matrix $\cal A$ is defined as 
\begin{align*}
\cal A :&= \frac{\rho_1}{\rho_2}\left[ \frac{2(w-1)(2-w)}{w^5}\kappa^2 - \frac{4d^2}{w^5}\right](X_2^{\top}X_2)^2 + \left[\frac{2\left( w-1\right)}{w^3}\kappa^2 - \frac{2d^2}{w^3} \right](X_1^{\top}X_1)^2 - 2 \frac{\sigma^2}{w^3}X_1^\top X_1.
\end{align*}
Using the estimate Fact \eqref{lem_minv}, we get that $\cal A$ is lower bounded as
\begin{align*}
\cal A \succeq&~ - \frac{4d^2}{w^5}n_1n_2 (\al_+(\rho_2)+\oo(1))^2 + \left[\frac{2\left( w-1\right)}{w^3}\kappa^2 - \frac{2d^2}{w^3} \right]n_1^2 (\al_-(\rho_1)-\oo(1))^2 \\
&~ - 2 \frac{\sigma^2}{w^3}n_1(\al_{+}(\rho_1)+\oo(1)) \succ 0,
\end{align*}
as long as
$$w> w_1:=1 +\frac{d^2}{\kappa^2}+ \frac{\sigma^2}{n_1\kappa^2}\frac{\al_{+}(\rho_1)+\oo(1)}{\al_{-}^2(\rho_1)} + \frac{2d^2}{\kappa^2}\frac{\rho_2(\al_+^2(\rho_2)+\oo(1))}{\rho_1\al_-^2(\rho_1) }.$$
Hence $h'(w)>0$ on $(w_1,\infty)$, %i.e. $h(w)$ is strictly increasing for $w>w_1$. This 
which gives $\hat w\le w_1$.

For the case $1/2\le w\le 1$, the proof is similar as above. Taking derivative of $h(w)$, we obtain that
\begin{align}
	h'(w) \le& \frac{\rho_1}{\rho_2} \left[2\left( w-1\right) \kappa^2\right]\cdot \tr\left[( w^2X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_2^{\top}X_2)^2\right] \nonumber \\
	& +  \left[2wd^2 +2w\left( w-1\right)(2w-1)\kappa^2\right]\cdot \tr\left[( w^2X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \nonumber\\
	& + \frac{\rho_1}{\rho_2}(2 w\sigma^2) \cdot \bigtr{(w^2X_1^{\top}X_1 + X_2^{\top}X_2)^{-2} X_2^\top X_2  }= \frac{\rho_1}{\rho_2} \bigtr{(w^2X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} \cal B} , \nonumber
\end{align}
where the matrix $\cal B$ is
$$\cal B= 2\left( w-1\right) \kappa^2  (X_2^{\top}X_2)^2+\frac{\rho_2}{\rho_1}\left[2wd^2 +2w\left( w-1\right)(2w-1)\kappa^2\right](X_1^{\top}X_1)^2 + 2 w\sigma^2 X_2^\top X_2 .$$
Using Fact \ref{lem_minv}, we get that $\cal B$ is upper bounded as
\begin{align*}
\cal B \preceq - 2(1-w)\kappa^2 n_2^2 (\al_-(\rho_2) -\oo(1))^2 +2w d^2 n_1n_2 (\al_+(\rho_1) +\oo(1))^2 + 2w\sigma^2 n_2 (\al_+(\rho_2)+\oo(1)) \prec 0,
\end{align*}
as long as
$$w< w_2:=1 -   \frac{d^2}{\kappa^2}\frac{\rho_1(\al_+(\rho_1) +\oo(1))^2}{\rho_2 \al_-^2(\rho_2) } -  \frac{\sigma^2}{n_2\kappa^2}\frac{\al_{+}(\rho_2)+\oo(1)}{\al_{-}^2(\rho_2)} .$$
Hence $h'(w)<0$ on $[0,w_2)$, %i.e. $h(w)$ is strictly decreasing for $w<w_2$. This 
which gives $\hat w\ge w_2$.

@@

Inserting \eqref{hatv_add1} into \eqref{eq_te_mtl_2task} and applying Lemma  \ref{largedeviation} to $(\beta_1-\hat v\beta_s)$, we get that w.h.p.,
\begin{align}
\te(\hat{\beta}_2^{\MTL})&=(1+\OO(\cal E))\cdot \left[d^2 + \OO\left(\cal E^2 \kappa^2\right)\right] \tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right]\nonumber\\
&+(1+\OO(\cal E))\cdot \sigma^2  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }. \label{1Liso}
\end{align}
In order to study the phenomenon of bias-variance trade-off, we need the bias term with $d^2$ and the variance term with $\sigma^2$ to be of the same order. Using Fact \eqref{lem_minv}, we have that
$$\tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \sim p,\quad \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} } \sim 1.$$
Hence we need to choose that $p d^2 \sim \sigma^2$. On the other hand, we want the error term $\cal E^2 \kappa^2$ to be much smaller than $d^2$, which leads to the condition $p^{-1+\e_0+4\e}\kappa^2  \ll d^2 \ll \kappa^2$. The above considerations lead to the choices of parameters in \eqref{choiceofpara}, under which we can simplify \eqref{1Liso} to \eqref{simple1}.
%\begin{align}
%\te(\hat{\beta}_2^{\MTL})&=(1+\OO(n^{-\e}))\cdot d^2 \tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \nonumber \\
%&+(1+\OO(n^{-\e}))\cdot \sigma^2  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }  \label{simple1}
%\end{align}
%w.h.p. for some constant $\e>0$.
\end{proof}

@@

\iffalse
\subsection{Task Similarity}\label{app_proof_31}

@@

\fi
